Self Referencing Sequences

### 2014-11-05 draft.tex
updating the determinism section

We claim that a self-referencing sequence is uniquely determined by the generating cycle and the starting number.

\begin{lemma}
Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
\end{lemma}

\begin{proof}
Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.

\begin{figure}{hb}
\label{fig:unequalsequences}
\center
\includegraphics[width=0.5\linewidth]{lemma1.pdf}
\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
\end{figure}

This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
\end{proof}

We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
\begin{quote}
\textbf{Algorithm:}\\
Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\

This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
\end{quote}

>>>>

We claim that a self-referencing sequence is uniquely determined by the generating cycle $C$ and the starting number $a \in C$. We will prove this by showing that if we are given the first number of the sequence, there is only one way to extend the sequence. In order to do this, we formalize the idea of breaking up a sequence into blocks.

\begin{definition}
A block $B$ is any maximal length subsequence of repeated numbers in a sequence. We denote by $B_i$ the $i$th block of the sequence.
\end{definition}

Here, \emph{maximal length} is a local property that means that the the block cannot be extended to include the numbers preceding or succeeding the block, as they will not be the repeated number. Thus, if a sequence $\{s_i\}$ is
\begin{align*}
1,3,3,3,1,1,1,3\dots \,
\end{align*}
the first block is $B_1 = \{1\}$, the second is $B_2 = \{3,3,3\}$, the third is $B_3 = \{1,1,1\}$, and so on.

It is obvious that a sequence can be specified by giving the numbers of the sequence individually, or by specifying all the blocks $B_i$. This is enough to make the following claim.

\begin{proposition}
A self-referencing sequence $s$ is uniquely specified by the generating cycle $C$, and the starting number $a \in C$. In other words, such an $s$ is guaranteed to exist and is unique.
\end{proposition}
\begin{proof} 
The proof is straightforward. The contents of block $B_i$ are determined by the value of $i$ and $s_i$. The value of $i$ determines which number is repeated throughout the block. 
\end{proof}
%
%\begin{lemma}
%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
%\end{lemma}
%
%\begin{proof}
%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
%
%\begin{figure}{hb}
%\label{fig:unequalsequences}
%\center
%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
%\end{figure}
%
%This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
%\end{proof}
%
%
%
%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
%\begin{quote}
%\textbf{Algorithm:}\\
%Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
%
%This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
%\end{quote}
